Telemetry Module for Solar Vehicle

### 2021-06-17 documentation/FinalPaper/paper.tex

\usepackage[pdftex]{graphicx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{float}
\usepackage{comment}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[style=ieee]{biblatex}
\addbibresource{paper.bib}
\usepackage{url}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}
\newcommand{\MYhref}[3][blue]{\href{#2}{\color{#1}{#3}}}

@@

\section{Design}
The overall system consists of three separate modules; the Solar Car Unit which communicates with the Support Vehicle Unit via a wireless RF connection, which in turn interacts with the Support Vehicle Laptop through a serial USB connection. The schematic for the system is outlined in Figure \ref{fig:schematic}. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/schematic.png}
    \caption{Overview of the system's components in block diagram form.}
    \label{fig:schematic}
\end{figure}
%

\section{Components}
\subsection{CAN bus}
The CAN network in the solar car enables the micro-controllers at each electrical unit in the car to send and receive data without a host computer. This is possible with the special structure of the CAN message.

@@

The frame uses an identifier to tell which messages goes to which unit. When two messages try to access the bus, the one with the highest priority, lowest ID, will be granted access to the bus, while the other message has to wait. After the ID comes the RTR bit which determines whether the frame remotely requests or actually sends data to a node. The control field, sometimes called DLC, holds the amount of bytes used in the following frame field containing the actual data. The rest of the fields in fig. \ref{fig:CANframe} serve error checking and coordination of communication on the bus.

The physical CAN bus itself consists of a differential pair of twisted wires, CAN High and CAN Low. These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively. A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.5V and the CAN Low wire is driven at 2.5V.

@@

An advantage of using the off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^\copyright$ \cite{shockburst}. Package structures contain a preamble, address, packet, data payload and CRC bytes segments as shown in Figure \ref{fig:shockburst}.

@@

We chose a SD card as a permanent local storage solution for the solar cars "black box" due to its ease of use, its interchangeability and since the Teensy 3.6 has an integrated SD card slot on its PCB. Furthermore, this solution makes it easy to increase the storage capacity to more than sufficient sizes with new SD card sizes exceeding 1 TiB.

\subsection{Microcontrollers} % Steffan
Finally, Teensy 3.6 and 4.0 microcontrollers are used as connection hubs in the Solar Car and Support Vehicle Units respectively. The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data. Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.

@@

Several of the key functionalities of the telemetry module are implemented in software. This together with the aim of designing a system which is easy to extend in the future makes the choice of software frameworks very important, since they affect reusability and maintainability of the developed code.

@@

\section{Verification and validation} % Victor
As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design. We undertook the software verification after finishing the prototype and software coding, reviewing the naming, functionality and commenting the code accordingly.

@@

\section{Discussion}
% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
% range test
In order to test the connection between the Solar Car and Support Vehicle modules we performed a range test in a suitable outdoor environment. In the end, we achieved a transmission range of around 160 m before the connection became too unreliable. This is obviously a big step down from the 1000 m claimed by the manufacturer. However, A video documenting the test can be viewed with \MYhref{https://youtu.be/-iwST3REn40}{this link}. 
% throughput limitations

>>>>

\usepackage[pdftex]{graphicx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{float}
\usepackage{comment}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[style=ieee]{biblatex}
\addbibresource{paper.bib}
\usepackage{url}
\usepackage[linkcolor = blue]{hyperref}
\newcommand{\MYhref}[3][blue]{\href{#2}{\color{#1}{#3}}}

@@

\section{Design}
The overall system consists of three separate modules; the Solar Car Unit (SCU) which is connected to the CAN bus and communicates with the support vehicle unit via a wireless RF connection. The Support Vehicle Unit (SVU) in turn interacts with the Support Vehicle laptop through a serial UART connection. The GUI program running on the laptop broadcasts the CAN bus message stream on a UDP port on the local network. A block diagram for the system is outlined in Fig. \ref{fig:schematic}. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/SystemSchematic.pdf}
    \caption{Overview of the system's components in block diagram form.}
    \label{fig:schematic}
\end{figure}
%

\section{Components}
\subsection{CAN bus}
The CAN network in the SCU enables the all the car's processing nodes to send and receive data without a host computer. This is possible thanks to the special structure of the CAN message.

@@

The frame uses an identifier to tell which messages goes to which unit. When two messages try to access the bus, the one with the highest priority (lowest ID) will be granted access to the bus, while the other message have to wait. After the ID comes the RTR bit which determines whether the frame remotely requests or actually sends data to a node. The control field, sometimes called DLC, holds the amount of bytes used in the following frame field containing the actual data. The rest of the fields in fig. \ref{fig:CANframe} serve error checking and coordination of communication on the bus.

The physical CAN bus itself consists of a differential pair of twisted wires, CAN High and CAN Low. These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively. A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.75V and the CAN Low wire is driven at 1.25V.

@@

An advantage of using the off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^\copyright$ \cite{shockburst}. Package structures contain a preamble, address, packet, data payload and CRC bytes segments as shown in Fig. \ref{fig:shockburst}. Only the data payload and address needs to be specified prior to sending a package; the preamble, packet and CRC segments are generated automatically and therefore fall beyond the scope of this project.

@@

We chose a SD card as a permanent local storage solution for the SCU's "black box" due to its ease of use and accesibility through the Teensy 3.6 integrated SD card slot. Furthermore, this solution scales very easily for larger storage capacity, with new SD card sizes exceeding 1 TiB.

\subsection{Microcontrollers} % Steffan
Finally, Teensy 3.6 and 4.0 microcontrollers are used as connection hubs in the SCU and SVU respectively. The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data. Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.

@@

Several of the key functionalities of the telemetry module are implemented in software. This, together with the aim of designing a system to be expanded upon in the future, makes the choice of software frameworks very important. Thus, an emphasis has been placed on the reusability and maintainability of the developed code.

@@

\section{Testing and Validation} % Victor
As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design. We undertook the software verification after finishing the prototype and software coding, reviewing the naming, functionality and commenting the code accordingly.

@@

%Finally, we were able to test all three modules

\section{Discussion}

% range test
In order to test the connection between the Solar Car and Support Vehicle modules we performed a range test in a suitable outdoor environment. In the end, we achieved a transmission range of around 160 m before the connection became too unreliable. A video documenting the test can be viewed with \MYhref{https://youtu.be/-iwST3REn40}{this link}. This is obviously a big step down from the 1000 m claimed by the manufacturer. However, considering the crude implementation of our hardware components, as well as the interference from buildings and other vehicles, that range was highly unattainable in the first place. Further enhancements in the hardware, such as a stable power supply for the RF modules, would likely improve the range beyond 400 m.

% throughput limitations
% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
